\section{Other Applications of Counterfactual Explanations}
\label{sec:otherapplications}

Here we refer the readers to other applications where counterfactual explanations are being used apart from explaining ML models:
\begin{enumerate}
    \item \emph{Anomaly and data-drift detection}: \citet{hinder-cfe-drift} propose to use CFEs to explain data drift. \citet{sulem-cfe-time-series-anomaly-detection} propose to use CFEs to explain anomalies in time-series datasets. \citet{ravi-cfe-anomaly-detection} wrote a survey on the explainability techniques for convolutional auto-encoders for anomaly detection of images. \citet{cfe_anomaly_detection} propose to use CFEs to explain anomaly detection when using autoencoders. \citet{antoran2021-cfe-for-uncertainty-estimation} use CFEs to find changes in a datapoint that would help a classifier have a higher confidence in its prediction. 

    \item \emph{Training dataset debugging}: \citet{debugging-ml-models-cfe} propose to use CFEs to debug ML models by diagnosing the behavior and using synthetic data to alter the decision boundaries. \citet{debugging-data-cfe2} propose to use CFEs to debug potentially mislabeled datasets. \citet{cfe-financial-risk-debugging} propose to use CFEs to detect bugs in financial models. \citet{debugging-data-cfe3-han} propose finding a minimal subset of training datapoints that are responsible for a particular prediction and hence can be used to debug training datasets. 
    
    \item \emph{Data augmentation}: \citet{CFE-earning-call-data-augment} propose to use CFEs to augment training data that is used to predict market volatility based on earning calls. 
    \citet{CFE-classimbalance} propose using CFEs to augment training data to tackle the class imbalance problem. 
    \citet{hasan-cfe-augmentation-proximity,rasouli-cfe-robustness-tabular} propose using CFEs for data augmentation of tabular datasets for increased robustness. 
    \citet{temraz-cfe-data-augmentation-climate-change} propose using CFEs to generate data points that can be used to train ML models that predict crop growth (afflicted by climate change). 
    
    \item \emph{Drug designing}: \citet{nguyen2021-cfe-dta} use CFEs to find changes in a drug and protein molecule that will increase their affinity for each other. They use multi-agent RL to this end. 
    
    \item \emph{ML model bias detection}: \citep{Ustun19:Actionable,cfe-model-bias1, cfe-model-bias2}. 
    
    \item \emph{Various applications}: \citet{mazzine-CFE-employment} propose to use CFEs in employment services to help job seekers get personalized advice for increasing their propensity for getting recommended for a job and to help the ML developers to detect potential bias and other issues in their ML model. \citet{Sadler-cfe-community-detection} propose to use CFEs for community detection in social networks. \citet{dimension_reduction_cfe} propose to use CFEs to understand interactive dimensionality reduction. \citet{student-performance-cfe} propose to use CFEs for providing actionable suggestions to improve student performance in a university course. \citet{Cong-CFE-KS-test} propose a CFE approach to explain why a test set fails the Kolmogorov-Smirnov test. \citet{Marchezini2022-dh-cfe-latent-mental-health} propose to use CFE for altering both observational and latent variables to reason about mental health. \citet{yao-cfe-evaluate-recsys} propose to use counterfactuals for evaluating the explanations for recommender systems. \citet{gupta-cfe-constraint-satisfaction-problem} use CFEs to propose changes to constraint satisfaction problems that have no solutions. \citet{cfe-for-entity-resolution} propose using CFEs to explain entity resolution models. \citet{artelt-cfe-model-adaptation} use CFEs to explain the differences between the learning of a pair of models. 
    \citet{CRASS-dataset-NLP} propose a new dataset, CRASS, to test reasoning and natural language understanding of LLMs. 
\end{enumerate}

There has been one case of real-world deployment of CFEs in a hiring platform, \href{https://hired.com/}{Hired}. \citet{nemirovsky-cfe-gan-hiring-application} use a GAN-based approach~\citep{nemirovsky-hired-people-cfe-countergan} to suggest changes in features like expected salary, years of experience, and skills to candidates in order to get them approved by the Hired Marketplace ML model. 

